---
title: "Power Analysis"
author: "Iris Kemp"
date: "August 14, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(pwr)
```

## Background for Pam
We're putting together a proposal for an experimental release study. Basically, we need to know whether we're actually going to be able to measure a statistically significant change in age structure and survival across the study groups.

The study set up is:

- Three release groups (early, normal, late) from our Glenwood facility.
- 100k fish per release group.
- Response variables: survival and age structure.

Glenwood survival averages 0.55% (see below table).

Age structure proportions for Glenwood returns: 

```{r}
glenwood <- data_frame(Year = 2007:2012, 
                       age2 = c(.0317, .085, .045, .0614, .10, .058), 
                       age3 = c(.7148, .5621, .6725, .7325, .5111, .6196), 
                       age4 = c(.2394, .3529, .2775, .2061, .3667, .3152), 
                       age5 = c(.0141, 0, .005, 0, .0222, .0072), 
                       survival = c(0.0071, 0.0058, 0.0074, 0.0053, 0.0019, 0.0052))

knitr::kable(glenwood)
```

(Q to follow up on: Does age structure differ between males and females? Beckman suggested splitting by sex. Assume no sex difference for now.)

<br>

## Power calculations

This is from one of our collaborators and I don't actually understand why he's doing what he's doing here. He's calculating average age of return based on the porportions of each age class that returns, and then using that single value to represent the returns and replicating over two groups and six years. But that doesn't really address whether the age structure (i.e., the distribution) is different, right? 

> This is just a quick cut at the problem, and only one way to approach it, but might give a sense of power.  First, I summed the proportion at each age x age for each year. 

| Year | Sum of (agexprop) |
|------|-------------------|
| 2007 |	3.236            |
| 2008 |	3.268            |
| 2009 |	3.243            |
| 2010 |	3.145            |
| 2011 |	3.311            |
| 2012 |	3.272            |

mean	3.246

stdev	0.056

> Then you can get a mean and stdev among years.  If you were to conduct an experiment with just two treatments (for simplicity) that ran for 6 years and had the following effect

| Age | prop(before) | prop(after) |
|-----|--------------|-------------|
| 2	  | 0.06         | 0.03        |
| 3	  | 0.64	       | 0.33        |
| 4	  | 0.29	       | 0.54        |
| 5	  | 0.01	       | 0.10        |
| Mean|	3.25	       | 3.71        |


> Here's a quick power analysis for two groups and 6 years with an effect size of 3.71-3.25 = 0.46; using the stdev from existing data.

> There are problems with this approach with a time series, and I'm sure other problems with it, but...

> pwr.1way(k=2, n=6, alpha=0.05, f = 0.46,  sigma = 0.056)

     Balanced one-way analysis of variance power calculation 

              k = 2
              n = 6
    effect.size = 0.46
      sig.level = 0.05
          power = 0.3026356

> NOTE: n is number in each group, total sample = 12, power = 0.302635637783944

<br> 

Would a chi-squared goodness of fit test be more appropriate?

We can average age structure over 2007-2012 and use that as null age structure (as Barry did above).

Assume sample size in each group is 150 adult returns. I don't know offhand how to do a chi-squared test for a null and two treatment groups so we'll just do null vs. alt for now. For alt age structure, we'll use year 2010 since it's the closest to null.

```{r}
null <- glenwood %>%
  select(., -Year, -survival) %>%
  summarise_all(mean) %>% 
  unlist(., use.names = FALSE)

y2010 <- glenwood[4,2:5] %>% unlist(., use.names = FALSE)

pwr.chisq.test(w = ES.w1(null, y2010), N = 150, df = (4-1), sig.level = 0.05)
```

So in this example we'd have about a 60% chance of rejecting null (i.e., concluding there was significant difference between age structures). 

Complicating factors: 

- My example only represents two treatment groups and one year, which is simpler than the actual experiment will be.

- We have no idea how age structure might change with the three treatments. Is there some way to simulate a bunch of different age structures? Maybe bounded by the observed data (this is only semi-reasonable, since the observed data only represent our normal releases - we don't have any early/late release data to compare)? 

- Consider observed survival data - ranges from 0.19% to 0.74%. So in a bad year, if we release 100k fish, we might only get 190 fish back. Assuming equal male/female ratios (not a particularly great assumption - it's probably more like 60:40 or even 70:30 depending on year), that'd be a sample size of 95.

- If survival varies wildly among treatment groups, we might end up with three totally different sample sizes of returns. This is probably pretty likely, based on experimental releases in other places.
