---
title: "Power Analysis"
author: "Iris Kemp"
date: "August 14, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

## Background
Three release groups (early, normal, late) from Glenwood - 100k fish per release group.
Response variables: survival and age structure.

Glenwood survival averages 0.55% (see below table).

Age structure percentages for Glenwood returns (per Barry spreadsheet): 

```{r}
glenwood <- data_frame(Year = 2007:2012, 
                       age2 = c(3.17, 8.5, 4.5, 6.14, 10, 5.8), 
                       age3 = c(71.48, 56.21, 67.25, 73.25, 51.11, 61.96), 
                       age4 = c(23.94, 35.29, 27.75, 20.61, 36.67, 31.52), 
                       age5 = c(1.41, 0, 0.5, 0, 2.22, 0.72), 
                       survival = c(0.71, 0.58, 0.74, 0.53, 0.19, 0.52))

knitr::kable(glenwood)
```

Does age structure differ between males and females? Beckman suggested splitting by sex.

<br>

## Barry's back-of-the-envelope

This is just a quick cut at the problem, and only one way to approach it, but might give a sense of power.  First, I summed the proportion at each age x age for each year. 

| Year | Sum of (agexprop) |
|------|-------------------|
| 2007 |	3.236            |
| 2008 |	3.268            |
| 2009 |	3.243            |
| 2010 |	3.145            |
| 2011 |	3.311            |
| 2012 |	3.272            |

mean	3.246

stdev	0.056

Then you can get a mean and stdev among years.  If you were to conduct an experiment with just two treatments (for simplicity) that ran for 6 years and had the following effect

| Age | prop(before) | prop(after) |
|-----|--------------|-------------|
| 2	  | 0.06         | 0.03        |
| 3	  | 0.64	       | 0.33        |
| 4	  | 0.29	       | 0.54        |
| 5	  | 0.01	       | 0.10        |
| Mean|	3.25	       | 3.71        |


Here's a quick power analysis for two groups and 6 years with an effect size of 3.71-3.25 = 0.46; using the stdev from existing data.

There are problems with this approach with a time series, and I'm sure other problems with it, but...

pwr.1way(k=2, n=6, alpha=0.05, f = 0.46,  sigma = 0.056)

     Balanced one-way analysis of variance power calculation 

              k = 2
              n = 6
    effect.size = 0.46
      sig.level = 0.05
          power = 0.3026356

NOTE: n is number in each group, total sample = 12, power = 0.302635637783944
